Reproducible Research: Peer Assessment 1
=============================================
```{r}
Sys.time()
```
This is my report for the assessment 1 of "Reproductive Research."

## Loading and preprocessing the data
I loaded the dataset, "Activity monitoring data." I saved the data as "raw.data."
```{r}
raw.data<-read.csv("activity.csv")
head(raw.data)
nrow(raw.data)
```
This raw data has many NA values and I made a data frame that does not have NAs.
```{r}
data.no.NA<-subset(raw.data,is.na(raw.data[,"steps"])==FALSE)
head(data.no.NA)
```
The date on which the mesurement was taken is;
```{r}
date.names<-levels(factor(raw.data$date))
date.names<-paste(substring(date.names,6,7),"/",substring(date.names,9,10),sep="")
head(date.names,3)
tail(date.names,3)
```
5-minute interval in which mesurement was taken is;
```{r}
int.names<-levels(as.factor(raw.data$interval))
head(int.names,3)
tail(int.names,3)
```
In the following, I used "ggplot2" so that I can make elegant graphics.
```{r,warning=FALSE}
library(ggplot2)
```

## What is mean total number of steps taken per day?
I saved the total number of steps taken per day to "tsd," ignoring the missing values.
```{r,results='markup'}
total.step.day<-tapply(data.no.NA$steps,data.no.NA$date,sum)
tsd<-data.frame(date=date.names,tsd=total.step.day)
head(tsd)
```
And I made a histogram with ggplot2.
```{r tsd}
g<-ggplot(tsd,aes(x=tsd))
g<-g+geom_histogram(binwidth=1000)
g+labs(list(title="The total number of steps taken per day",
               x="Total number of steps per each day",y="Frequency"))
```
The mean and median of total number of steps taken per day is in this summary.
```{r}
summary(tsd)
```
To see only mean and median, I made objects, meanA and medianA.
```{r}
#only mean
(meanA<-mean(tsd$tsd,na.rm=TRUE))
#only median
(medianA<-median(tsd$tsd,na.rm=TRUE))
```

## What is the average daily activity pattern?
I calculated the average number of steps taken for each intaval, with ignoring missing values. The values are averaged across all days.
```{r}
mean.int<-data.frame(interval=as.numeric(int.names),
                     mean=as.vector(tapply(data.no.NA$steps,
                                            data.no.NA$interval,mean)))
```
I made a time siries plot of the average number of steps taken vs. the 5-minutes interval.
```{r daily activity pattern}
q<-ggplot(mean.int,aes(x=interval,y=mean))+geom_line()
q+labs(list(title="average daily activity pattern",x="interval",
            y="the average number of steps taken for each intaval"))
```
The interval contains the maximum averaged number of steps is;
```{r}
sort.data<-mean.int[order(mean.int$mean,decreasing=TRUE),]
sort.data[1,"interval"]
```

## Imputing missing values
Because NAs sometimes introduce  bias into analysis of data, I have to filling the misisng values in the dataset. The number of missing values in the dataset is;
```{r}
nrow(subset(raw.data,is.na(raw.data[,"steps"])==TRUE))
```
For estimating the NA values, I used the arithmetic mean number of means of the day and the interval. For example, I explain about the first row of raw.data:
```{r}
raw.data[1,]
```
The mean number of steps of 10/2 is;
```{r}
mean.day<-data.frame(date=date.names,
                       mean=as.vector(tapply(data.no.NA$steps,
                                             data.no.NA$date,mean)))
mean.day[is.na(mean.day)]<-0
mean.day[1,]
```
the average number of steps taken for each intaval, with ignoring missing values;
```{r}
mean.int[is.na(mean.int)]<-0
mean.int[1,]
```
So, the NA value in the first row will be raplaced by the estimated number;
```{r}
(mean.day[1,2]+mean.int[1,2])/2
```
In the same way, 
```{r}
replaced<-raw.data

for (i in 1:nrow(replaced)){
  if(is.na(replaced[i,"steps"])==TRUE){
    mean.today<-subset(mean.day,
                       mean.day[,"date"]==
                         paste(substring(replaced[i,"date"],6,7),"/",
                               substring(replaced[i,"date"],9,10),sep=""))
    mean.the.int<-subset(mean.int,
                         mean.int[,"interval"]==replaced[i,"interval"])
    replaced[i,"steps"]<-(mean.today[1,2]+mean.the.int[1,2])/2
  }
}
head(replaced)
```
I use "replaced" in the following analysis. This dataset is a new dataset that is the same as the original dataset but with the missing cells filled in.

The total number of steps taken per day is;
```{r}
tsd2<-tapply(replaced$steps,replaced$date,sum)
tsd2<-data.frame(date.names,tsd2)
head(tsd2)
```
Here is the histogram of the new dataset;
```{r,tsd2}
p<-ggplot(tsd2,aes(x=tsd2))
p<-p+geom_histogram(binwidth=1000)
p+labs(list(title="The total number of steps taken per day  (with missing data filled in)",
            x="Total number of steps per each day",y="Frequency"))
```
And I calculated the mean and median;
```{r}
summary(tsd2)
meanB<-mean(tsd2$tsd2,na.rm=TRUE)       #only mean
medianB<-median(tsd2$tsd2,na.rm=TRUE)   #only median
```
To compare tsd and tsd2, I made multiple histograms in a screen.
```{r, warning=FALSE}
library(gridExtra)
g<-g+labs(list(title="The total number of steps taken per day",
               x="Total number of steps per each day",y="Frequency"))
p<-p+labs(list(title="The total number of steps taken per day (with missing data filled in)",
            x="Total number of steps per each day",y="Frequency"))
grid.arrange(g,p,nrow=2)
```
And I overayed multiple histograms.
```{r}
pg<-ggplot(data=tsd,aes(x=tsd))
pg<-pg+geom_histogram(binwidth=1000)
pg<-pg+geom_histogram(data=tsd2,aes(x=tsd2),binwidth=1000,alpha=0.5)
pg+labs(list(title="The difference of total number of steps taken per day with or without NAs",x="Total number of steps per each day",y="Frequency"))
```
The difference between tsd and tsd2 is in the gray zone.
This zone makes the differences of mean and median number.
```{r}
meanA-meanB
medianA-medianB
```

## Are there differences in activity patterns between weekdays and weekends?
First, I made a factor which has two levels-"weekday" and "weekend."
I set English as the language R prints with, and and added new column to the data frame "replaced."
```{r}
#print "Date" in English
Sys.setlocale("LC_TIME","English")
#to see weekdays
replaced$weekdays<-weekdays(as.Date(raw.data$date),abbreviate=TRUE)
#convert into weekday,weekend
replaced$weekdays[replaced$weekdays=="Sun"|
                    replaced$weekdays=="Sat"]<-"weekend"
replaced$weekdays[replaced$weekdays!="weekend"]<-"weekday"
```
Next, I calculated the mean numbers of steps taken, averaged across all weekday s or weekend days.
```{r}
end.data<-subset(replaced,replaced$weekdays=="weekend")
day.data<-subset(replaced,replaced$weekdays=="weekday")

end.int<-as.vector(tapply(end.data$steps,factor(end.data$interval),mean))
day.int<-as.vector(tapply(day.data$steps,factor(day.data$interval),mean))

week.int<-data.frame(interval=as.numeric(int.names),weekend=end.int,weekday=day.int)
head(week.int)
```
Last, I made a panel plot conteining a time series plot, using ggplot2 and gridExtra.
```{r}
a<-ggplot(week.int,aes(x=interval,y=weekend))+geom_line(col="blue")
b<-ggplot(week.int,aes(x=interval,y=weekday))+geom_line(col="blue")
a<-a+labs(list(title="average daily activity pattern on weekends",
               x="interval",y="average number of steps"))
b<-b+labs(list(title="average daily activity pattern on weekdays",
               x="interval",y="average number of steps"))
grid.arrange(a,b,nrow=2)
```


## Citation

Data;
https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip

Explanations for this assessment in GitHub;
https://github.com/rdpeng/RepData_PeerAssessment1.git

Tools of analysis;
R Core Team (2013). R: A language and environment for statistical
  computing. R Foundation for Statistical Computing, Vienna, Austria. URL
  http://www.R-project.org/.

H. Wickham. ggplot2: elegant graphics for data analysis. Springer New
  York, 2009. URL http://had.co.nz/ggplot2/book

Baptiste Auguie (2012). gridExtra: functions in Grid graphics. R package
  version 0.9.1. http://CRAN.R-project.org/package=gridExtra

